Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures

### 2018-09-05 supplement.tex
Add section/plots "benefits of neg for pos design" to supplement

>>>>

\section{Immediate benefits of negative design for positive design}

Already previous work on RNA Design, e.g. INFO-RNA and Incarnation, observed that---briefly put---positive design benefits from negative design. Here, more concretely, we only \emph{demonstrate} this effect (for small examples), where we consider \Def{positive design} as  
designing RNA sequences that fold with high probability $Pr(R|S)$ into the given target structure $R$; and \Def{negative design}, as generating one or many sequences that have a low energy for $R$.

For a double-stem structure, we produce a uniform and a weighted Boltzmann sampling, each of 1000 sequenes. For each sampled sequence $S$, we determine the Turner energy $E(S,R)$ as well as the
%difference $E_\Delta(S,R)$ of $E(S,R)$ to the ensemble energy $E_{\text{ens}}(S)$ of $S$ 
probability $Pr(R|S)$ of the structure $R$ in the ensemble of sequence $S$ 
(where the energies are calculated in the Turner model using RNAfold).

Note that low energy $E(S|R)$ is (non-linearly) correlated with high probability $Pr(S|R)$, the probability to see $S$ in the Boltzmann distribution of sequences folding into $R$. Recall that the probability $Pr(R|S)$ can be calculated as $\exp(-E(S,R)/RT)/Z(S)$,
where $Z(S)=\sum_{R'} \exp(-E(S,R')/RT)$. 
%
Consequently, Figure~\ref{fig:negforpos} ultimately illustrates the dependency between $Pr(S|R)$ and $Pr(R|S)$ for uniform samples and low energy samples (where the latter can be efficiently generated due to our Boltzmann sampling strategy).

% and high energy difference $E_\Delta(S,R)=E_{\text{ens}}(S)-E(S,R)$, with high probability $\Pr(R|S)$.
% 
% To see the latter, recall that the \Def{ensemble energy $E_{\text{Ens}}(S)$} is defined as
% $$
% E_{\text{ens}}(S) = -RT \ln( Z(S) ),
% $$ 
% based on the partition function $Z(S)=\sum_{R'} \exp(-E(S,R')/RT)$
% such that 
% $$
% E(S,R) - E_{\text{ens}}(S) 
% =
% -RT \ln Pr(R|S),
% $$  
% which can be derived by term transformation 
% \begin{align*}
% -RT \ln P(R|S) &= -RT \ln [ exp(-E(S,R)/RT) / Z(S) ] \\&= -RT \ln (exp(-E(S,R)/RT)) - (-RT\ln Z(S))
% \\&=  E(S,R) - E_{\text{Ens}}(S). 
% \end{align*}

\begin{figure}
\textbf{A}~\includegraphicstop[width=0.48\textwidth]{Figs/Plots/doublestem_weight1}%
\textbf{B}~\includegraphicstop[width=0.48\textwidth]{Figs/Plots/doublestem_weight500}
\caption{Comparision of (A) Uniform sampling to (B) Boltzmann sampling with weight $\pi=500$. For given structure $R$, we plot the energies $E(S,R)$ of sampled sequences against the probabilities $Pr(R|S)$; the former is controlled by Boltzmann sampling, while the latter is a common positive design objective.}
\label{fig:negforpos}
\end{figure}
